\normalsize{A general approach and tool to evaluate the effectiveness of no-take marine reserves}

### 2017-07-21 Manuscript.Rmd
Added conclusions and gave some formatting

Juan Carlos Villaseñor-Derbez^1\*^, Caio Faro^1^, Melaina Wright^1^, Jael Martínez^1^, Alvin Suárez^2^, Stuart Fulton^2^, Sean Fitzgerald^1^, Fiorenza Micheli^3^, Gavin McDonald^4^, Jorge Torre-Cosio^2^, Christopher Costello^1,4^ \

@@

Marine reserves are often implemented to preserve habitat and recover overfished stocks. In Mexico, recent regulations allow fishers to propose legally-recognized no-take marine reserves, established by the fishery management agency (CONAPESCA) for a period of time. While the regulation includes instructions to request their implementation, no guidelines are provided for monitoring, evaluation, or renewal. Since 2012, 35 reserves have been implemented, often managed by the communities and supported by local NGOs. Yearly biological monitorings are conducted in  reserves and their control sites. To date, the evaluation of the reserves relies solely on biological data, ignoring socioeconomic and governance dimensions. In absence of a standardized methodology, existing data is analyzed in different ways, hindering the ability to compare results across communities. We developed a framework to evaluate the effectiveness of the reserves by matching 7 commonly stated management objectives to a set of 9 biophysical, 5 socioeconomic, and 14 governance indicators. This framework includes a standardized survey to collect governance and socioeconomic data. Biophysical indicators are evaluated with a Difference-in-Difference analysis, estimating the net effect of the reserve through time. Linear regression models are fitted to socioeconomic indicators through time, testing for the difference in trends before and after the implementation of the reserve. Governance indicators are analyzed based on literature, identifying common governance structures and their associated effectiveness. To make the framework accessible to fishers and decision makers, and allow replication of results, we developed an open source application. The results produced by the application contain a report for a general audience, and a technical report that can inform the renewal and decision-making process.

@@

Based on these seven objectives, we determined a set of indicators to evaluate the effectiveness. The list of indicators was built through a review of scientific literature, where we identified indicators that were used to measure similar objectives. The indicators were reviewed at a workshop by members of Mexican fishery management agencies and non-government organizations. Later, these were also presented to fishers from the Ensenada Fishing Cooperative (*S.C.P.P. Ensenada*), which  provided input. Our final list of indicators includes, at least, those identified in review works such as @lester_2009-Ks or @woodcock_2017-Wm.

@@

MAREA was developed in R Studio [@rcore_2017] using the shiny framework [@shiny_2017]. This provides the tools to build interactive web applications, hosted on an open server. MAREA can be accessed at [turfeffect.shinyapps.io/marea/](turfeffect.shinyapps.io/marea/). In order to use the tool, a user shoudl have previously identified the management objectives of the reserve and the data. MAREA automatically selects the appropriate inicators, which can be modified by users to adjust to their interests. MAREA produces two types of results, aimed for different audiences.

@@

The scorecard is produced with functions from the Shinydashboard package [@shinydashboard_2017]. The technical report is produced by a parameterized Rmarkdown document [@rmarkdown_2017] processed by the knitr package [@knitr_2017].

@@

```{r map, fig.cap = "General location of Isla Natividad (left) and map of the island (right). The La Plana / Las Cuevas Marine Reserve polygon is indicated in gray, and the approximate location of the La Dulce / Babencho control site is shown with no fill."}
proj <- CRS("+proj=longlat +datum=NAD27")

@@

For govenance data, we constructed the indicators based on our knowledge and familiarity of the area, as well as prototype interviews performed to liders of the fishing cooperative.

@@

@@

The results shown here intend to highlight the relevance and utility of the framework and tool, which automate the analysis and make it replicable. First, we present the scorecard generated by MAREA to provide an overview of the state of the reserve. Then, we provide (mdoified for style) a set of figures generated by MAREA, showing time-series for biophysical and socioeconomic indicators as well as tables containing the magnitude and significance of the $\beta_2$ and $\beta_1$. A table with the governance indicators is also provided and discussed. While presenting each result, we aknowledge possible limitations that the framework and tool might have and provide possible solutions.

@@

The scorecard shows that, overall, the reserve has a good performance achieving a general score of 63.6% of positive indicators. All category-level scores were also high, with values of 66.7%, 60%, and 75% positive indicators for Biophysical, Socioeconomic and Governance, respectively.

@@

@@

The authors wish to thank the fishers from Isla Natividad, who gathered the data used in this study for over 10 years, and the fishers from El Rosario, who helped us validate our survey and framework. We also thank Olivier Deschenes and Andrew Plantinga, who provided valuable input to design the model that evaluates the biophysical indicators.

## Funding

# Other plots

>>>>

Juan Carlos Villaseñor-Derbez^1\*^, Caio Faro^1^, Melaina Wright^1^, Jael Martínez^1^, Alvin Suárez^2^, Stuart Fulton^2^, Sean Fitzgerald^1^, Gavin McDonald^4^, Jorge Torre-Cosio^2^, Fiorenza Micheli^3^, Christopher Costello^1,4^ \

@@

Marine reserves are often implemented to preserve habitat and recover overfished stocks.

To date, the evaluation of the reserves relies solely on biological data, ignoring socioeconomic and governance dimensions. In absence of a standardized methodology, existing data is analyzed in different ways, hindering the ability to compare results across communities. We developed a framework to evaluate the effectiveness of the reserves by matching 7 commonly stated management objectives to a set of 9 biophysical, 5 socioeconomic, and 14 governance indicators. This framework includes a standardized survey to collect governance and socioeconomic data. Biophysical indicators are evaluated with a Difference-in-Difference analysis, estimating the net effect of the reserve through time. Linear regression models are fitted to socioeconomic indicators through time, testing for the difference in trends before and after the implementation of the reserve. Governance indicators are analyzed based on literature, identifying common governance structures and their associated effectiveness. To make the framework accessible to fishers and decision makers, and allow replication of results, we developed an open source application. The results produced by the application contain a report for a general audience, and a technical report that can inform the renewal and decision-making process.

@@

Based on these seven objectives, we determined a set of indicators to evaluate the effectiveness. The list of indicators was built through a review of scientific literature, where we identified indicators that were used to measure similar objectives. A first filter eliminated indicators for which baseline data does not exist, or for which data aquisition was expensive or outside of the general scope. The initial list of indicators was reviewed at a workshop with participation of members from Mexican fishery management agencies and non-government organizations. Later, these were also presented to fishers from the Ensenada Fishing Cooperative (*S.C.P.P. Ensenada*), which provided input. Our final list of indicators includes, at least, those identified in review works such as @lester_2009-Ks or @woodcock_2017-Wm.

@@

MAREA was developed in R Studio [@rcore_2017] using the shiny framework [@shiny_2017]. This provides the tools to build interactive web applications, hosted on an open server. MAREA can be accessed at [turfeffect.shinyapps.io/marea](turfeffect.shinyapps.io/marea/). While the original version was developed in spanish (aimed mainly at Latinamerican countries), all its content can be translated by a translation tool availabe within the environment.

MAREA is designed into a 6-step process. The first part provides an introduction to the app, and provides the user with an overview of the process. Then, the user must select management objectives, and MAREA will automatically select appropriate indicators, based on **Table 2**. Then, the user can load the data, using standard \*.csv formatting; sample datasets are provided in the environment. Once data have been loaded, MAREA identifies all reserves in the data, and lets the user select the reserve to be evaluated. At this point, the user can also specify the year of implementation of the reserve, as well as select objective species that are of particular management interest. Before presenting the results MAREA provides the user with a section to confirm that all the decisions taken before are correct. Finally, the user is taken to the results, which are represented in two ways.

@@

The scorecard is produced with functions from the Shinydashboard package [@shinydashboard_2017]. The technical report is produced by a parameterized Rmarkdown document [@rmarkdown_2017] processed by the knitr package [@knitr_2017]. Another feature of MAREA is that the user can choose to share the data. Once the technical report is downloaded, information on the reserve, its management objectives, and all uploaded data is saved into a central repository. These data can be accessed at any time by any person interested in aquiring them.

@@

```{r map, fig.cap = "General location of Isla Natividad (left) and map of the island (right). The marine reserve polygon is indicated in gray, and the approximate location of the control site is shown with no fill."}
proj <- CRS("+proj=longlat +datum=NAD27")

@@

For govenance data, we constructed the database based on our knowledge of the area, as well as prototype interviews performed to liders of the fishing cooperative.

@@

The results shown here intend to highlight the relevance and utility of the framework and tool, which automate the analysis and make it replicable. First, we present the scorecard generated by MAREA to provide an overview of the state of the reserve. Then, we provide (mdoified for style) a set of figures generated by MAREA, showing time-series for biophysical and socioeconomic indicators as well as tables containing the magnitude and significance of the $\beta_2$ and $\beta_1$ coefficients. A table with the governance indicators is also provided and discussed.

The scorecard (**Fig. 4**) shows that, overall, the reserve has a good performance achieving a general score of 63.6% of positive indicators. All category-level scores were also high, with values of 66.7%, 60%, and 75% positive indicators for Biophysical, Socioeconomic and Governance, respectively.

@@

@@

@@

One of the greatest challenges in management measuring the extent to which objectives have been met. The present framework provides a simple and straightforward way to align management objectives with performance indicators. We aknowledge that these 25 indicators might not fully describe a particular reserve. However, we believe that these provide a starting point to perform the evaluation, and urge decision-makers, managers, and users to include other indicators (*e.g.* larval dispersal or connectivity) that are relevant to their reserve.

The proposed methodologies, specially the way in which biophysical indicators are evaluated, provide valuable information for managers. The analysis isolates the net effect of the reserve, providing a propper measure of reserve effectiveness. We agree there is room for improvement in the way in which socioeconomic and governance data are analized. However, we believe that providing a unifying platform where all can be analyzed and comprehensively presented represents a valuable step towards effective management.

The main value of MAREA is that it provides a free, simple, and replicable way to perform rigurous analysis. The tool can easily be used by fishers, NGO members, and managers in government agencies, providing transparency of the analysis and results. The way in which results are presented allow this information to be interpreted by a wider audience. The scorecard is easily understandable by experts and non-experts, and can be used as an effective tool for communicating the results of yearly evaluations. On the other side, the technical report can serve as a tool for managers and scientists to rapidly communicate information at a more technical level.

While the first release of MAREAis now available, it will continue to be developed and maintained. This will incorporate new features, and enhance current ones, aiming to improve user experience and expand the scope of the analysis. Yet, we believe that this first release represents a major step towards effective management of marine reserves.

@@

The authors wish to thank the fishers from Isla Natividad, who gathered the data used in this study, and the fishers from El Rosario, who helped us validate our survey and framework. Input provided by the participants in the workshop "" was very valuable. We also thank Olivier Deschenes and Andrew Plantinga, who provided valuable input to design the model that evaluates the biophysical indicators.
